\pkg{ggenealogy}: An \proglang{R} Package for Visualizing Genealogical Data

### 2016-11-06 article.tex
replication script update

\begin{CodeChunk}
\begin{CodeInput}
R> install.packages("ggenealogy")
R> library("ggenealogy")
R> data("sbGeneal")
R> str(sbGeneal)
\end{CodeInput}
\begin{CodeOutput}
'data.frame':	390 obs. of  5 variables:
 $ child       : chr  "5601T" "Adams" "A.K." "A.K. (Harrow)" ...
 $ year        : num  1981 1948 1910 1912 1968 ...
 $ yield       : int  NA 2734 NA 2665 NA 2981 2887 2817 NA NA ...
 $ year.imputed: logi  TRUE FALSE TRUE FALSE FALSE FALSE ...
 $ parent      : chr  "Hutcheson" "Dunfield" NA "A.K." ...
\end{CodeOutput}
\end{CodeChunk}
\subsection{Academic genealogy of statisticians}
The \pkg{ggenealogy} package also comes with an academic genealogy of statisticians; this dataset is in the form of a data frame with 8165 rows and six columns. To develop this later dataset, we contacted the \citealt{mgp}, a web-based database for the genealogy of academic mathematicians. This database, which currently contains almost 200,000 entries, is a service of the North Dakota State University Department of Mathematics and the American Mathematical Society. The Mathematics Genealogy Project contact provided us a Structured Query Language (SQL) export, and we used PostgreSQL to query the database (\citealt{psql}).

@@

\begin{CodeChunk}
\begin{CodeInput}
R> data("statGeneal")
R> dim(statGeneal)
\end{CodeInput}
\begin{CodeOutput}
[1] 8165    6
\end{CodeOutput}
\begin{CodeInput}
R> colnames(statGeneal)
\end{CodeInput}
\begin{CodeOutput}
[1] "child"   "parent"  "year"    "country" "school"  "thesis"
\end{CodeOutput}
\end{CodeChunk}

@@

However, some \pkg{ggenealogy} plots also make use of quantitative variable values associated with individuals in the genealogy. For these plots, the input data frame should also contain a third quantitative variable column. Each case in this quantitative variable column can be of type numeric, type NA, or type ``". In both example data files, this column is labeled ``year,". However, we would like to emphasize that the name of quantitative variable column to be plotted is user-specified. For instance, should the data of interest contain time-related values on a shorter scale than years (such as days or months), then the column could be named ``day" or ``month". This flexibilitiy allows for more diverse datasets to easily be examined with the package.

@@

\begin{CodeChunk}
\begin{CodeInput}
R> pathTN <- getPath("Tokyo", "Narow", sbIG, sbGeneal)
R> pathTN
\end{CodeInput}
\begin{CodeOutput}
$pathVertices
[1] "Tokyo"    "Volstate" "Jackson"  "R66-873"  "Narow"   
$yearVertices
[1] "1907"   "1942"   "1954.5" "1971.5" "1985" 
\end{CodeOutput}
\end{CodeChunk}

@@

Next, we can run the same set of functions on a different pair of varieties. First, a call to the \pkg{ggenealogy} function \code{getYear()} indicates that variety \code{Bedford} was released in 1978 and variety \code{Zane} in 1985.

\begin{CodeChunk}
\begin{CodeInput}
R> getYear("Bedford", sbGeneal)
\end{CodeInput}
\begin{CodeOutput}
[1] 1978
\end{CodeOutput}
\begin{CodeInput}
R> getYear("Zane", sbGeneal)
\end{CodeInput}
\begin{CodeOutput}
[1] 1985
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{Code}
R> plotPathOnAll(pathTN, sbGeneal, sbIG, bin = 1:3, pathEdgeCol = "red",
+    nodeSize = 2.5, pathNodeSize = 4) + ggplot2::theme(axis.text = ggplot2::
+    element_text(size = 12), axis.title = ggplot2::element_text(size = 12))
\end{Code}

@@

\begin{Code}
R> plotPathOnAll(pathTN, sbGeneal, sbIG, bin = 1:6, pathEdgeCol =
+    "seagreen2", nodeSize = 1, pathNodeSize = 3)
\end{Code}

@@

@@

\begin{CodeChunk}
\begin{CodeInput}
R> statIG <- dfToIG(statGeneal)
R> pathCB <- getPath("David Cox", "Petra Buzkova", statIG, statGeneal,
+    isDirected = FALSE)
R> plotPath(pathCB, fontFace = 4) + ggplot2::theme(axis.text =
+    ggplot2::element_text(size = 10), axis.title =
+    ggplot2::element_text(size = 10)) + + ggplot2::scale_x_continuous(expand
+    = c(.1, .2))
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> plotPathOnAll(pathCB, statGeneal, statIG, bin = 1:200) +
+    ggplot2::theme(axis.text = ggplot2::element_text(size = 8), axis.title =
+    ggplot2::element_text(size = 8)) + ggplot2::scale_x_continuous
+    (expand = c(.1, .2))
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> plotPathOnAll(pathCB, statGeneal, statIG, bin = 1:200, nodeSize = .5,
+    pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol = "lightgray") +
+    ggplot2::theme(axis.text = ggplot2::element_text(size = 8), axis.title =
+    ggplot2::element_text(size = 8)) + ggplot2::scale_x_continuous
+    (expand = c(.1, .2))
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> plotPathOnAll(pathCB, statGeneal, statIG, bin = 1:200, nodeSize = .5,
+    pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol = "lightgray",
+    animate = TRUE)
\end{CodeInput}
\end{CodeChunk}

>>>>

\begin{CodeChunk}
\begin{CodeInput}
R> install.packages("ggenealogy")
R> library("ggenealogy")
R> data("sbGeneal")
R> str(sbGeneal)
\end{CodeInput}
\begin{CodeOutput}
'data.frame':	390 obs. of  5 variables:
 $ child       : chr  "5601T" "Adams" "A.K." "A.K. (Harrow)" ...
 $ date        : num  1981 1948 1910 1912 1968 ...
 $ yield       : int  NA 2734 NA 2665 NA 2981 2887 2817 NA NA ...
 $ date.imputed: logi  TRUE FALSE TRUE FALSE FALSE FALSE ...
 $ parent      : chr  "Hutcheson" "Dunfield" NA "A.K." ...
\end{CodeOutput}
\end{CodeChunk}
\subsection{Academic genealogy of statisticians}
The \pkg{ggenealogy} package also comes with an academic genealogy of statisticians; this dataset is in the form of a data frame with 8165 rows and six columns. To develop this later dataset, we contacted the \citealt{mgp}, a web-based database for the genealogy of academic mathematicians. This database, which currently contains almost 200,000 entries, is a service of the North Dakota State University Department of Mathematics and the American Mathematical Society. The Mathematics Genealogy Project contact provided us a Structured Query Language (SQL) export, and we used PostgreSQL to query the database (\citealt{psql}).

@@

\begin{CodeChunk}
\begin{CodeInput}
R> data("statGeneal")
R> dim(statGeneal)
\end{CodeInput}
\begin{CodeOutput}
[1] 8165    6
\end{CodeOutput}
\begin{CodeInput}
R> colnames(statGeneal)
\end{CodeInput}
\begin{CodeOutput}
[1] "child"   "parent"  "date"    "country" "school"  "thesis" 
\end{CodeOutput}
\end{CodeChunk}

@@

However, some \pkg{ggenealogy} plots also make use of quantitative variable values associated with individuals in the genealogy. For these plots, the input data frame should also contain a third quantitative variable column. Each case in this quantitative variable column can be of type numeric, type NA, or type ``". In both example data files, this column is labeled ``date," and each case in that column can be of type numeric, type NA, or type ``". At this point, any \pkg{ggenealogy} plot can be used.

@@

\begin{CodeChunk}
\begin{CodeInput}
R> pathTN <- getPath("Tokyo", "Narow", sbIG, sbGeneal)
R> pathTN
\end{CodeInput}
\begin{CodeOutput}
$pathVertices
[1] "Tokyo"    "Volstate" "Jackson"  "R66-873"  "Narow"   
$dateVertices
[1] "1907"   "1942"   "1954.5" "1971.5" "1985"  
\end{CodeOutput}
\end{CodeChunk}

@@

Next, we can run the same set of functions on a different pair of varieties. First, a call to the \pkg{ggenealogy} function \code{getDate()} indicates that variety \code{Bedford} was released in 1978 and variety \code{Zane} in 1985.

\begin{CodeChunk}
\begin{CodeInput}
R> getDate("Bedford", sbGeneal)
\end{CodeInput}
\begin{CodeOutput}
[1] 1978
\end{CodeOutput}
\begin{CodeInput}
R> getDate("Zane", sbGeneal)
\end{CodeInput}
\begin{CodeOutput}
[1] 1985
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{Code}
R> plotPathOnAll(pathTN, sbGeneal, sbIG, bin = 3, pathEdgeCol = "red",
+    nodeSize = 2.5, pathNodeSize = 4) + ggplot2::theme(axis.text = ggplot2::
+    element_text(size = 12), axis.title = ggplot2::element_text(size = 12))
\end{Code}

@@

\begin{Code}
R> plotPathOnAll(pathTN, sbGeneal, sbIG, bin = 6, pathEdgeCol =
+    "seagreen2", nodeSize = 1, pathNodeSize = 3) + ggplot2::xlab("Year")
\end{Code}

@@

Notice from Figure~\ref{fig:plotTNBin3} that the default horizontal axis label for the \texttt{plotPath()} method has a value of ``Date". Given that the ``date" variable in this example dataset is on the timescale of years, we wanted to change the default value of the horizontal axis label to ``Year". We did this in the code above for Figure~\ref{fig:plotTNBin6} by appending the \texttt{ggplot2::xlab()} function.

@@

\begin{CodeChunk}
\begin{CodeInput}
R> statIG <- dfToIG(statGeneal)
R> pathCB <- getPath("David Cox", "Petra Buzkova", statIG, statGeneal,
+    isDirected = FALSE)
R> plotPath(pathCB, fontFace = 4) + ggplot2::xlab("Year") +
+    ggplot2::theme(axis.text = ggplot2::element_text(size = 10), axis.title =
+    ggplot2::element_text(size = 10)) + + ggplot2::scale_x_continuous(expand
+    = c(.1, .2))
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> plotPathOnAll(pathCB, statGeneal, statIG, bin = 200)
+    ggplot2::theme(axis.text = ggplot2::element_text(size = 8), axis.title =
+    ggplot2::element_text(size = 8)) + ggplot2::scale_x_continuous
+    (expand = c(.1, .2)) + ggplot2::xlab("Year")
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> plotPathOnAll(pathCB, statGeneal, statIG, bin = 200, nodeSize = .5,
+    pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol = "lightgray") +
+    ggplot2::theme(axis.text = ggplot2::element_text(size = 8), axis.title =
+    ggplot2::element_text(size = 8)) + ggplot2::scale_x_continuous
+    (expand = c(.1, .2)) + ggplot2::xlab("Year")
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> plotPathOnAll(pathCB, statGeneal, statIG, bin = 200, nodeSize = .5,
+    pathNodeSize = 2.5, nodeCol = "darkgray", edgeCol = "lightgray",
+    animate = TRUE)
\end{CodeInput}
\end{CodeChunk}
